\section{Conclusion}
\label{sec:conclusion}

This paper turned a ninth-century theory of recognition into a five-clause engineering specification for steering the verbalizable workspace of frozen language models, and subjected every clause to pre-registered, gate-audited, adversarially reviewed experiments. The confirmed mechanisms are concrete: instructed loadability that scales with model size and vanishes under pruning and distillation; band content legible only in band coordinates, with the articulation gradient shown to be model-intrinsic; entropy-gated writes that steer within a $\pm0.5$-nat freedom budget at six independent-stream seeds with a sign-consistent advantage over rate-matched and prefill controls; a cross-plant calibration recipe, amplitude inversely proportional to lens transport strength, confirmed monotone at four seeds and consistent across two model families over 1.5 orders of amplitude; and an untrained associative-memory write path that reproduces the analytic path at the metric's resolution. The registered negatives are equally concrete: gated steering yields legibility and budgeted placement, not raw surfacing throughput, and single contrastive directions purchase no alignment improvement on an already-aligned model under the registered criteria.

Three lines of future work are registered in the ledger. Training the associative store against steering-success targets, to test whether learning improves on cold recall. Extending readback verification past matched corpora, where it currently fails pooled generalisation. And scaling the interventional programme to the 27B plant, where loadability is strong and the doctrine's predictions sharpen. The research-loop architecture that produced these results, expected-free-energy selection over registered menus with dual-closure gates and isolated adversarial review, is released in full and is offered, together with the companion circuit-discovery agent \cite{Sathish2026ACD} and refusal-symmetry analysis \cite{Sathish2026Refusal}, as a template for rigorous interpretability research under small compute budgets.
