\label{sec:conclusion}

\subsection*{Summary}

This work operationalizes Pratyabhij\~n\=a's doctrine of reflexive awareness as an engineering specification
for language-model workspace steering, and provides an honest map of what the doctrine enables and what it does not.

\textbf{What the doctrine enables:}
(1) Workspace-band structure replicates across three model families and two sizes; instructed loadability scales
with model size and collapses under pruning/distillation. (2) Band-targeted lenses are necessary for readback;
final-target lenses are structurally blind to intermediate content. (3) Capped residual writes into the band steer
downstream behavior and preserve model freedom (entropy within $\pm 0.5$ nat budget) under sampling decoding.
(4) Event-gated timing (entropy-based sphura\d{t}\d{t}\=a detection) delivers full steering within budget,
confirmed across six independent-stream seeds (lift 0.30--0.35, sign-consistent alignment $p \approx 0.016$).
(5) The steering method transfers to a second plant via a calibration recipe where amplitude $\propto$ inverse
lens transport strength, confirmed at four independent-stream seeds with monotone dose response in each plant's
active range. (6) The research program ran as expected-free-energy selection over nine registered menus, with
twenty-six cycles and sixteen adversarial reviews catching real errors before merge.

\textbf{What the doctrine does not deliver:}
(1) Gated steering does not maximize concept surface-rate (CSR 0.16 vs continuous 0.96); it is a legibility-efficiency
tradeoff, not a raw-lift maximizer. (2) Naive contrastive direction extraction for refusal/truthfulness does not improve
downstream benchmarks on an already-aligned model (AdvBench ASR 0.25 $\to$ 0.25, TruthfulQA truthfulness 0.697 $\to$ 0.680).
Alignment improvements at the activation level are not automatic and may require training-time methods.
(3) The doctrine is fundamentally about workspace \emph{transparency} and reflexive re-cognition, not alignment assurance.
Practitioners should not expect band-steering to solve alignment alone; it is a tool for interpretability and internal-model
visibility, valuable in its own right.

\subsection*{Impact and Next Steps}

Pratyabhij\~n\=a's integration of \emph{what}, \emph{where}, \emph{when}, and \emph{how} supplies a missing theory layer
for interpretability.
The band-targeted lens, the sphura\d{t}\d{t}\=a timing gate, and the amplitude calibration recipe are now available tools
for steering open models.
The research loop's self-correction (through adversarial review and gate discipline) demonstrates that innovation
and rigor are not opposed; they co-enable each other.

Future work includes training the CittaStore against steering-success targets (the PWM write path is now integrated and
verified in L20 under a cold, untrained store; whether training beats cold initialization is open), cross-episode
continuity (anusa\d{m}dh\=ana), and the relationship between J-space and the broader global workspace.
The library (prabodha v1.0.0) and the research ledger are now public; practitioners can reproduce, extend, and refute
each claim.
