\label{sec:discussion}

\subsection*{Honest Negatives and Scope Clarifications}

The $\geq 0.15$ schedule margins (anticipated steering speedup under gating) did not confirm;
we observe sign consistency instead.
The \=agama readback acceptance test, recalibrated against behavioral hits and held-out-tested at
a fixed setting, achieves balanced accuracy 0.64 against a pre-registered 0.6 (gate L14-readback),
a margin inside its own binomial CI.
When pooling three corpora at $n=120$ (gate L15-readback, L16-corpus), the accuracy regresses to 0.59.
The readback signal is weak---not an acceptance test, but a directional hint.

\textbf{L21 findings refine the doctrine's scope:}
(1) Gated steering achieves full behavior-steering within the freedom budget (Finding 3), but concept
surface-rate ($0.16$ gated vs $0.96$ continuous) reveals that gating is a legibility-efficiency tradeoff,
not a raw-lift maximizer (Finding 8, gate L21-baselines). The doctrine is not ``writes will make concepts
surface at frequency $X$,'' but rather ``writes via gated moments preserve model autonomy while enabling
reflexive re-cognition.'' This distinction reframes band-steering as a governance mechanism for workspace
transparency, not a throughput optimizer.
(2) Naive contrastive alignment steering (extracting refusal or truthfulness directions from small labeled
sets) does not improve AdvBench ASR (baseline 0.25 $\to$ steered 0.25, gate L21-jailbreak) or TruthfulQA
truthfulness-proxy (baseline 0.697 $\to$ steered 0.680, gate L21-truthful). This is an honest negative: the
steering infrastructure does not unlock hidden alignment gains at the activation level. Alignment improvements
at scale appear to require training-time or inference-time techniques (filtering, ensemble voting, supervised
fine-tuning) outside the scope of activation steering (Finding 9).

\subsection*{Limitations: Metrics, Scale, and Experimental Design}

\textbf{Behavioral metric proxies:} CSR (concept surface-rate), ASR (attack success rate), and truthfulness-proxy
(string-overlap with reference answers) are heuristic-based. LLM-judge evaluation on the same tasks would strengthen
all claims, especially jailbreak and truthfulness findings (gate L21 notes this in the deviations field).
Our refusal-rate and ASR calculations use keyword matching; adversarial-robustness evaluation with human labeling
would be more robust. Truthfulness via string-overlap is notoriously loose; semantic similarity scoring or full
TruthfulQA evaluation would tighten confidence intervals.

\textbf{Model scale:} All confirm-tier claims (core steering, recipe transfer, dose response) derive from Qwen3-4B
or Nemotron-Mini-4B (both $\lesssim 5$B parameters). The L1b finding (loadability scales with size) suggests
27B models should show stronger band signals, but this was not carried forward systematically into the L21 alignment
work. Scaling the contrastive direction experiments (jailbreak, truthfulness) to larger models remains open.

\textbf{Single-configuration contrastive steering:} L21 jailbreak and truthfulness experiments extract directions from
a single labeled set (10 and 15 examples, respectively) and apply them in the same band. Extracting multiple directions
(e.g., from different datasets or labelers) and testing their composition or conflict remains unexplored.
The generalization of these directions to prompts outside the source domain is not tested.

\textbf{Corpus and concept selection:} The concept corpus (25 instructed concepts on Qwen3-4B) is hand-selected and
small. Scaling to larger, more diverse concept sets would strengthen generalization claims. The L16-corpus finding
that steering-lift varies with corpus style suggests that corpus diversity itself is a confound that should be factored
out with careful experimental design (e.g., randomized concept-to-corpus assignment).

\textbf{Band identification methodology:} The band is identified via CKA clustering of attribution matrices across layers,
with an assumption that the band's boundary is sharp. Fuzzy band boundaries or overlapping band structure are not explored;
band-membership uncertainty (e.g., soft membership) could be incorporated into future steering designs.

The single-seed fragility program reviewers flagged in narrative-past corpus style was resolved into a mechanism:
the calibration recipe's amplitude term has a \emph{corpus} axis (stub difficulty) alongside its model axis (lens strength).
A fixed amplitude across corpora was the real gap (gate L16-corpus).
Once amplitude is calibrated per corpus, steering lift generalizes (descriptive-scene: 0.30/0.20/0.25 at $\alpha=0.1$;
narrative-past: 0.43/0.28/0.23 at $\alpha=0.2$, 3/3 seeds).
However, the commitment-flash reading (our implementation of sphura\d{t}\d{t}\=a) remains directionally weaker than
entropy gating, not refuted.

The trained-bridge comparator, comparing PWM's CittaStore write against prabodha's analytic band-exit write,
was executed in L20: the PWM stack is now integrated on the target hardware and the CittaStore write path runs
end-to-end on the frozen model. Under a \emph{cold} (untrained) store initialization---populated with no steering
patterns---the recall path steers within the entropy budget on all three seeds and is comparable to the analytic
baseline (exact match on 2/3 seeds; on the third the store differs by a single concept-hit, in its own favour,
which fails the strict $\le0.05$ equivalence criterion). Because the store was never trained, whether training it
against steering-success targets improves over cold initialization remains the open follow-up; the infrastructure
to test it is now live (gate \texttt{gate\_L20\_confirm.json}).

Three modalities remain deliberately unconverged:
\begin{enumerate}
\item {\bf Anusa\d{m}dh\=ana} (cross-episode continuity): the steering recipe is intra-episode;
      testing whether episode boundaries induce a reset of workspace loading is future work.
\item {\bf Modality confound}: all prompts are text-in, tokens-out; multimodal workspace structure is unexplored.
\item {\bf W-space beyond J-space}: the global workspace (GWT) may be larger than the $J$-space defined by
      our J-space band identification; the relationship remains open.
\end{enumerate}

\subsection*{Reproducibility and Program Discipline}

The research loop caught and corrected four classes of errors on the record:
\begin{enumerate}
\item Cost model miscalibration (L3 resolved by review \#2).
\item Replay bug in the ledger (L8 state mismatch discovered in review \#7, corrected in L9).
\item Sampling-seed correlation (the \emph{stream fix}, applied retroactively to all L8--L19 gates).
\item Staleness violations (gate-to-gate propagation; now enforced by ledger linter).
\end{enumerate}

Every gate JSON includes pre-registration, amendment log, evidence (raw numbers, p-values, CIs), tier (smoke/screen/confirm),
and verdict. The entire program consumed $\sim 18$ GPU-hours on GB10, co-resident with training jobs,
with contention events logged in every gate.
