\label{sec:intro}

\subsection*{Motivation and the Pratyabhij\~n\=a Parallel}

The interpretability community possesses an actuator---the Jacobian lens---but lacks a doctrine of \emph{use}.
Pratyabhi\~jñ\=a, a ninth-century non-dual philosophy, supplies that doctrine.
Its core insight is that reflexive awareness (\emph{vimarśa}) = the supreme Word (\emph{par\=a v\=ak}) defines consciousness
(ĪPK 1.5.13).
We read this as an engineering specification for language models:
if the global workspace is linguistic verbalizable reflexivity, then a doctrine of use follows
immediately---{\em what} to write (verbalizable content codes), {\em where} (the workspace band in band coordinates),
{\em when} (at moments of uncommitted awareness), {\em how to verify} (the workspace re-reads what was offered).

\subsection*{Related Work and Positioning}

Global Workspace Theory (GWT; Baars, 1988; Dehaene, 2014) predicts that consciousness arises from a cognitive
workspace where multiple processes broadcast information to a global state. The interpretability community has
built tools for reading and writing this workspace---Sparse Autoencoders (SAEs), activation steering (RepE, Zou et al.,
2023), Contrastive Activation Analysis (CAA; Rimsky et al., 2024), function vectors, and logit lenses.
However, prior work focuses on \emph{what can be done with these tools}, not on \emph{a doctrine of responsible use}.

Activation steering methods share key mechanics with our approach: both use direction vectors extracted from model
activations, both target specific layers, and both rely on linear combinations of residual stream values.
CAA (Rimsky et al., 2024) extracts contrastive directions from labeled pairs (harmful/benign, refused/compliant)
and applies them to steering, similar to our L21 jailbreak and truthfulness experiments.
RepE (Zou et al., 2023) and function vectors use regression and principal components to find steering directions.
Our contribution is orthogonal: we supply a (1) timing doctrine (entropy-gated writes, \emph{sphura\d{t}\d{t}\=a}-detection),
(2) a readback protocol for verifying workspace engagement (band-targeted lenses, \emph{\=agama}-recognition),
(3) a calibration recipe for cross-plant transfer (amplitude $\propto$ 1/lens-strength), and (4) a philosophy-as-engineering-specification
methodology that makes the steering doctrine falsifiable.
L21's honest negative on contrastive alignment steering (Finding 9) is evidence that steering alone does not solve alignment---
a finding that complements rather than contradicts prior CAA work, which tested on smaller refusal margins and different models.

Pratyabhij\~n\=a supplies what prior steering work lacks: it operationalizes the \emph{what}, \emph{where}, \emph{when}, and \emph{how}
as falsifiable engineering clauses, each with a distinct signal and failure mode, and grounds the whole program in a rigorous
theory of verbalizable workspace structure.

\subsection*{Research Questions and Scope}

This work tests whether Pratyabhij\~n\=a's doctrine survives contact with frozen-decoder LLMs,
and characterizes what mechanisms it enables versus what it does not. Specifically:
\begin{enumerate}
\item Does the workspace-band structure replicate across model families and sizes?
\item Can a band's content be read only by a lens targeted at the band itself?
\item Do capped residual writes steer behavior within a freedom budget?
\item Does event-gated timing deliver steering efficiency?
\item Does the calibration recipe transfer across plants?
\item Does band-targeted concept steering surface information as efficiently as continuous flooding,
      or is the tradeoff a feature, not a bug---transparency at efficiency cost?
\item Can naive contrastive direction extraction improve alignment/refusal on an already-aligned model?
\item Can the research loop itself run as expected-free-energy selection?
\end{enumerate}

\subsection*{Contributions}

This paper presents nine key findings---six that confirm the Pratyabhij\~n\=a doctrine's core mechanisms,
and three that honestly characterize its scope limits. Every finding is verified at confirm or screen tier with
pre-registration, multi-seed replication where applicable, and a complete replayable research ledger from which
every number can be traced to a committed gate JSON. The contribution is framed as an honest map: what the
doctrine enables (workspace transparency, reflexive re-cognition, efficiency within budget) and what it does
not (raw concept-surface throughput, downstream alignment gains from naive contrastive steering alone). This
honesty is the paper's defining strength---overclaiming steering as an alignment shortcut would be technically
cleaner, but false.
