\label{sec:method}

\subsection*{The Doctrine of Use: Five Engineering Clauses}

Pratyabhij\~n\=a identifies five aspects of the workspace: {\em what} (verbalizable content),
{\em where} (the band), {\em when} (uncommitted moments), {\em how} (re-reading), and {\em freedom} (sv\=atantrya).
We operationalize each.

\subsubsection*{What: Verbalizable Concept Codes}

A concept code is a direction in the residual stream such that adding it increases the predicted probability
of tokens associated with a verbalizable concept (e.g., ``fire'', ``metal'', ``river'').
Such codes are constructed by fitting a linear regression on a small corpus of in-distribution examples
and are used without further tuning on the test corpus.

\subsubsection*{Where: The Workspace Band in Band Coordinates}

The workspace band is the final layers (typically 24--31 on a 32-layer model) where output-logit correlation
is high.
Writes are applied to the residual stream at these depths, not to earlier layers.
Writing at the wrong depth produces minimal steering effect.

\subsubsection*{When: Sphura\d{t}\d{t}\=a (Commitment Flash)}

At each decode step, the model's predictive distribution over the next token defines the ``committed'' output
of the current step.
Entropy measurements discriminate: high entropy = uncommitted moment; low entropy = moment of commitment.
Sphura\d{t}\d{t}\=a gates writes to high-entropy moments.
Contrast: rate-matched baselines write on a fixed schedule; greedy-decoding baselines write at all steps.

\subsubsection*{How to Verify: \=Agama Re-cognition (Readback via Band-Targeted Lenses)}

After writing to the band, the model reads back its own state via the residual stream.
We verify this re-reading via a band-targeted lens: if the write succeeded, the lens readout should reflect
the attempted concept.
A final-target lens (fitted to output logits) will miss band content and produce false negatives.

\subsubsection*{Budget: Sv\=atantrya (Freedom and Cost)}

Every write reduces the model's entropy: the local cost at the write moment is large ($-1.9$ to $-2.1$ nats
at-position) and decoding-independent.
The trajectory-average cost depends on regime: $+0.82$ nats under greedy decoding, $-0.13$ under sampling.
We register trajectory-average under sampling as the budget currency and set a $\pm 0.5$-nat cap.

\subsubsection*{The Program as Active-Inference Loop}

The research program runs 24 cycles over three experiment menus, selecting the next experiment via
expected-free-energy (EFE) maximization.
Beliefs replay from the program's own gate verdicts; observations are gate verdicts and safety margins;
spend is tracked in a JSONL ledger.
Eleven adversarial reviews (isolated agents, gate+contract briefs only) caught four classes of errors:
cost-model miscalibration, replay bugs, sampling-seed correlations, and staleness violations.
Every honest negative is recorded in the ledger and the commit history.
